---
title: "Divvy Fun"
output:
  html_document:
    df_print: paged
---
Goal here is to do some geographic mapping of divvy routes using their data. Using all of 2017 data and Q3/Q4 routes. Found here:
https://www.divvybikes.com/system-data

You will see that there's way too much data for me to run all routes and it will be reflected in the code

Step 1: load data and packages 
```{r}
#install what you need and load (for now). I don't think I use all of these packages but just load them - no ones gonna care! I don't remember where i pulled this snippit from! 


library(dplyr)
require(maps)
require(ggmap)
library(raster)
library(sp)
library(maptools)
library(lubridate)
gpclibPermit()
library(maptools)
library(mapdata)
library(ggmap)
library(geosphere)
library(maps)
library(ggplot2)



#pulling in the data - you'll need to useyour own file paths below. Could I do this all in 1 line? Probably.
div_stationQ3Q4 = read.csv("C:\\Users\\Brian\\Downloads\\Divvy_Trips_2017_Q3Q4\\Divvy_Stations_2017_Q3Q4.csv", header = TRUE)
div_tripsQ4 = read.csv("C:\\Users\\Brian\\Downloads\\Divvy_Trips_2017_Q3Q4\\Divvy_Trips_2017_Q4.csv", header = TRUE)
div_tripsQ3 = read.csv("C:\\Users\\Brian\\Downloads\\Divvy_Trips_2017_Q3Q4\\Divvy_Trips_2017_Q3.csv", header = TRUE)
div_tripsQ2 = read.csv("C:\\Users\\Brian\\Downloads\\Divvy_Trips_2017_Q1Q2\\Divvy_Trips_2017_Q2.csv", header = TRUE)
div_tripsQ1 = read.csv("C:\\Users\\Brian\\Downloads\\Divvy_Trips_2017_Q1Q2\\Divvy_Trips_2017_Q1.csv", header = TRUE)

```

Step 2: Transform the dang data cause it's never how you want it and it never will be so dplyr your way through life.

```{r}
#taking the quarterly data, adding a column to announce it, making it into one file and exporting.
#why? well I thought I could plot it all at once and my cowworkers wanted the tranformed single yearly dataset

Q1<-div_tripsQ1 %>% mutate(Quarter="Q1")
Q2<-div_tripsQ2 %>% mutate(Quarter="Q2")
Q3<-div_tripsQ3 %>% mutate(Quarter="Q3")
Q4<-div_tripsQ4 %>% mutate(Quarter="Q4")

Final<-bind_rows(Q1,Q2,Q3,Q4)
##write.csv(Final, file = "final_2017.csv")

#always check out your tables. Final lol - its never final

#joining the stations and trip info with the from_station_id form the trips data and the id from the stations

Q4_2017_coord <- dplyr::left_join(div_tripsQ4, div_stationQ3Q4,c('from_station_id'='id'))

#dropping the useless stuff and renaming to it make a little bit of sense
Q4<-Q4_2017_coord %>% dplyr::select(-c(city,name,X))
names(Q4)[names(Q4) == 'latitude'] <- 'from_lat'
names(Q4)[names(Q4) == 'longitude'] <- 'from_lon'
names(Q4)[names(Q4) == 'dpcapacity'] <- 'from_dpcap'

#joining the to_station_ids to station id's like before and renaming
Q4_2017_coord <- dplyr::left_join(Q4, div_stationQ3Q4,c('to_station_id'='id'))
Q4final<-Q4_2017_coord %>% select(-c(city,name,X))
names(Q4final)[names(Q4final) == 'latitude'] <- 'to_lat'
names(Q4final)[names(Q4final) == 'longitude'] <- 'to_lon'
names(Q4final)[names(Q4final) == 'dpcapacity'] <- 'to_dpcap'
head(Final)
```

Step 3: build the maps! Now I have to 

a. create a goodlooking map -google maps are pretty!

b. plot the stations - I want to know the relative sizes of stations too

c. map the routes from station to station for a quarter, or a random sample of 10,000 rows because I did this last minute and have a real job with real responsibilties and this is all just for fun.

```{r}
#building our map of Chicago. Centering it on the mean of my longitude and latitude of divvy stations
#i know - clever!

map_of_Chicago <- get_map(location = c(lon = mean(div_stationQ3Q4$longitude), lat = mean(div_stationQ3Q4$latitude)), zoom = 12, maptype = "roadmap", scale = 2) 

#overlaying a bunch of crap
#Geom_segment are the trip coordinates mapped - only using 10,000 datapoints because my PC is a potato
#geom_point are the stations - the size of the points are based on the number of bike ports available
map<-ggmap(map_of_Chicago)+ geom_segment(aes(x = from_lon , y = from_lat, xend = to_lon, yend = to_lat, colour = "segment", alpha=0.1), data =sample_n(Q4final,10000,replace=TRUE))+ geom_point(data = div_stationQ3Q4, aes(x = longitude, y = latitude, fill = "red", alpha = 0.8), size = div_stationQ3Q4$dpcapacity/14, shape = 21) + guides(fill=FALSE, alpha=FALSE, size=FALSE)

```

